---
title: "Explanation Analysis"
output:  
  html_document:
    self_contained: yes
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)
library(dplyr)
library(scales)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


# Analysis of subjects' explanations

## What subjects would have done if they hadn't been forced to make a choice

```{r}
library(readr)
explan_data <- read_delim("subjects_explanations_analysis_main.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)


colnames(explan_data)[3]  <- "different_explanation"        # subjects who gave an explanation that seemed to have nothing to do with visibility 
colnames(explan_data)[4]  <- "no_explanation"               # subjects who either gave no explanation for what they would do, or their explanation was unclear
colnames(explan_data)[5]  <- "misunderstood_scenario"       # subjects who wrote something that indicates that they didn't understand the scenario correctly
colnames(explan_data)[6]  <- "equally_likely"               # subjects who wrote explanations saying that 50:50 would have been their preferred answer
colnames(explan_data)[7]  <- "absence_because_visible"      # subjects who inferred the absence of the latent effect because it should have been visible
colnames(explan_data)[8]  <- "absence_because_description"  # subjects who inferred the absence of the latent effect for any other reason 
colnames(explan_data)[9]  <- "inferred_absence"             # overall number of subjects who inferred the absence of the latent feature 


# "nm" below means "not meet"; expressing the the criterion is not fulfilled

explan_data$different_explanation[explan_data$different_explanation == 1] <- "different_explanation"
explan_data$different_explanation[is.na(explan_data$different_explanation)] <- "nm"

explan_data$different_explanation <- factor(explan_data$different_explanation, levels = c("nm", "different_explanation"), 
                                     labels = c("nm", "different explanation"))

explan_data$no_explanation[explan_data$no_explanation == 1] <- "no_explan"
explan_data$no_explanation[is.na(explan_data$no_explanation)] <- "nm"

explan_data$no_explanation <- factor(explan_data$no_explanation, levels = c("nm", "no_explan"), 
                                     labels = c("nm", "no explanation or unclear"))

explan_data$misunderstood_scenario[explan_data$misunderstood_scenario == 1] <- "misunderstood"
explan_data$misunderstood_scenario[is.na(explan_data$misunderstood_scenario)] <- "nm"

explan_data$misunderstood_scenario <- factor(explan_data$misunderstood_scenario, levels = c("nm", "misunderstood"), 
                                     labels = c("nm", "misunderstood scenario"))


explan_data$equally_likely[explan_data$equally_likely == 1] <- "equally_likely"
explan_data$equally_likely[is.na(explan_data$equally_likely)] <- "nm"


explan_data$equally_likely <- factor(explan_data$equally_likely, levels = c("nm", "equally_likely"), 
                                     labels = c("nm", "equally likely"))



explan_data$absence_because_visible[explan_data$absence_because_visible == 1] <- "yes"
explan_data$absence_because_visible[is.na(explan_data$absence_because_visible)] <- "nm"


explan_data$absence_because_visible <- factor(explan_data$absence_because_visible, levels = c("nm", "yes"), 
                                     labels = c("nm", "absence because it should be visible"))



explan_data$absence_because_description[explan_data$absence_because_description == 1] <- "yes"
explan_data$absence_because_description[is.na(explan_data$absence_because_description)] <- "nm"


explan_data$absence_because_description <- factor(explan_data$absence_because_description, levels = c("nm", "yes"), 
                                     labels = c("nm", "absence given the description"))



explan_data$inferred_absence[explan_data$inferred_absence == 1] <- "yes"
explan_data$inferred_absence[is.na(explan_data$inferred_absence)] <- "nm"

explan_data$inferred_absence <- factor(explan_data$inferred_absence, levels = c("nm", "yes"),
                                       labels = c("nm", "inferred absence"))


```


## subjects giving no explanation or unclear explanation 

```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, no_explanation) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))


plotdata_between

plotdata_sub <- subset(plotdata_between, no_explanation == "no explanation or unclear")


```



```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Features,
           y = pct,
           fill = no_explanation)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Inferred state",
       x = "Features",
       title = "Subjects' inference about latent feature (overall)")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
#ggsave("absence_overall.svg",width=6,height=5)
```

```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[4]), n = c(40, 40), alternative = "two.sided", correct = F)
```

## Subjects clearly stating they'd prefer to answer 50:50


```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, equally_likely) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_between

plotdata_sub <- subset(plotdata_between, equally_likely == "equally likely")


```


```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Features,
           y = pct,
           fill = equally_likely)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1", labels = c("nm", "'equally likely option'")) +
  labs(y = "Percentage", 
       fill = "Categorization of \nsubjects' explanations",
       x = "Features")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
#ggsave("preferred_answer.svg",width=6,height=5)
```

```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[4]), n = c(40, 40), alternative = "greater", correct = F)
```


## Subjects who inferred the absence of the latent feature

Because of its visibility: 

```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, Knowledge, absence_because_visible) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_between

plotdata_sub <- subset(plotdata_between, absence_because_visible == "absence because it should be visible")


```



```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Features,
           y = pct,
           fill = absence_because_visible)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Inferred state",
       x = "Features",
       title = "Subjects' inference about latent feature")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
#ggsave("absence_because_visible.svg",width=6,height=5)
```


```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[4]), n = c(40, 40), alternative = "less", correct = F)
```


Inferred absence for any other reason: 

```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, absence_because_description) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_between

plotdata_sub <- subset(plotdata_between, absence_because_description == "absence given the description")


```


```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Features,
           y = pct,
           fill = absence_because_description)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Inferred state",
       x = "Features",
       title = "Subjects' inference about latent feature 2")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
#ggsave("absence_because_description.svg",width=6,height=5)
```

```{r}
prop.test(x = c(0, plotdata$n[3]), n = c(40, 40), alternative = "less", correct = F)
```


Inferred Absence of latent feature overall: 

```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, inferred_absence) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_between

plotdata_sub <- subset(plotdata_between, inferred_absence == "inferred absence")


```


```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Features,
           y = pct,
           fill = inferred_absence)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Inferred state",
       x = "Features",
       title = "Subjects' inference about latent feature (overall)")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
#ggsave("absence_overall.svg",width=6,height=5)
```


```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[4]), n = c(40, 40), alternative = "less", correct = F)
```


## Subjects giving a different kind of explanation (i.e., explanations that have nothing to do with visibility)

```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, different_explanation) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))


plotdata_between

plotdata_sub <- subset(plotdata_between, different_explanation == "different_explanation")
```



```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Features,
           y = pct,
           fill = different_explanation)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Inferred state",
       x = "Features",
       title = "Subjects' inference about latent feature (overall)")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
#ggsave("absence_overall.svg",width=6,height=5)
```


```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[4]), n = c(40, 40), alternative = "two.sided", correct = F)
```
