---
title: "Analysis"
output:  
  html_document:
    self_contained: yes
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)
library(dplyr)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_csv("exp_data.csv")
```


# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


# Data preperation

```{r, echo = FALSE}
# reorder factor 

# to create chart, the data must be in long format and only contain the relevant dependent variables

# 1. make a subset with only the relevant dvs

tdata_long <- tdata


tdata_long$scale_orient <- factor(tdata_long$scale_orient, levels = c("narrow_left", "narrow_right"), 
                               labels = c("narrow_left", "narrow_right"))



# Recode ratings so that high values represent broad scope and low values represent narrow scope

dat_narrow_left <- subset(tdata_long, scale_orient == "narrow_left")
dat_narrow_left$rating_rec <- (dat_narrow_left$DV_rating * 1)

dat_narrow_right <- subset(tdata_long, scale_orient == "narrow_right")
dat_narrow_right$rating_rec <- (1 - dat_narrow_right$DV_rating)

tdata_long <- rbind(dat_narrow_left, dat_narrow_right)


tdata_long$rating_rec <- factor(tdata_long$rating_rec, levels = c(1, 0), labels = c("broad", "narrow"))


library(stringr)
Features <- str_split_fixed(tdata_long$Scenario, "_", 2)[,1]
Knowledge <- str_split_fixed(tdata_long$Scenario, "_", 2)[,2]

tdata_long$Features <- Features
tdata_long$Knowledge <- Knowledge


table(tdata_long$Features, tdata_long$Knowledge)

tdata_long$Features <- factor(tdata_long$Features, levels = c("spearNet", "featherTooth"), labels = c("feature diagnosability: \nboth similar", "feature diagnosability: \nlatent feature harder"))

tdata_long$Knowledge <- factor(tdata_long$Knowledge, levels = c("notKnow", "noInfo"), labels = c("'you don't know'", "no information"))

```
# Analyses

```{r}
tdata_main <- subset(tdata_long, Knowledge == "'you don't know'")
tdata_supp <- subset(tdata_long, Knowledge == "no information")
```


```{r}
counts <- tdata_main %>%
  group_by(Features, rating_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts
```


```{r}

counts$category <- factor(counts$rating_rec, levels = c("unbiased", "narrow", "broad"), labels = c("unbiased", "narrow l.s.", "broad l.s."))

```



Get proportion CIs for different categories in each diagnosability condition:

```{r}
library(PropCIs)
library(DescTools)
library(purrr)

counts_spear <- subset(counts, Features == "feature diagnosability: \nboth similar")

counts_feather <- subset(counts, Features == "feature diagnosability: \nlatent feature harder")


(MultinomCI(counts_spear$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_1)

  
(MultinomCI(counts_feather$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_2)
  



ci_low <- c(selection_ci_1[,2], selection_ci_2[,2])

ci_up <- c(selection_ci_1[,3], selection_ci_2[,3])


```




```{r}
plotdata <- counts

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = category,
           y = pct,
           fill = Features)) +
  facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 3.5,
            position = position_dodge(width = 1),
            vjust = -5.5) +
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "none", axis.title = element_text(size = 20), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13),strip.text.x = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections_between.pdf",width=6,height=5)
ggsave("categories_main.svg",width=7,height=5)
ggsave("categories_main.pdf",width=7,height=5)
```



Tests against chance:

```{r}
(t2 <- binom.test(plotdata$n[2], 40))

(t4 <- binom.test(plotdata$n[4], 40))

(ci_low <- c(t2$conf.int[1], t4$conf.int[1]))

(ci_up <- c(t2$conf.int[2], t4$conf.int[2]))

(p_values <- c(t2$p.value, t4$p.value))

```

Test if proportion of narrow latent scope biases is different between conditions: 

```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[4]), n = c(40, 40), alternative = "two.sided", correct = F)
```

## Analysis with additional test query condition

The additional between-subjects condition was one in which the unobserved feature wasn't mentioned at all in the test query. In what follows, the graphs from above will be recreated and the additional condition will be included.


```{r}
counts <- tdata_long %>%
  group_by(Knowledge, Features, rating_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts
```


```{r}

counts$category <- factor(counts$rating_rec, levels = c("unbiased", "narrow", "broad"), labels = c("unbiased", "narrow l.s.", "broad l.s."))
counts
```



Get proportion CIs for different categories in each diagnosability condition:

```{r}
library(PropCIs)
library(DescTools)
library(purrr)

counts_dontknow_spear <- subset(counts, Knowledge == "'you don't know'" & Features == "feature diagnosability: \nboth similar")
counts_dontknow_feather <- subset(counts, Knowledge == "'you don't know'" & Features == "feature diagnosability: \nlatent feature harder")

counts_noinf_spear <- subset(counts, Knowledge == "no information" & Features == "feature diagnosability: \nboth similar")
counts_noinf_feather <- subset(counts, Knowledge == "no information" & Features == "feature diagnosability: \nlatent feature harder")


(MultinomCI(counts_dontknow_spear$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_1)

  
(MultinomCI(counts_dontknow_feather$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_2)
  

(MultinomCI(counts_noinf_spear$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_3)

(MultinomCI(counts_noinf_feather$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_4)





ci_low <- c(selection_ci_1[,2], selection_ci_2[,2], selection_ci_3[,2], selection_ci_4[,2])

ci_up <- c(selection_ci_1[,3], selection_ci_2[,3], selection_ci_3[,3], selection_ci_4[,3])


```




```{r}
plotdata <- counts

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = category,
           y = pct,
           fill = Features)) +
  facet_grid(Knowledge ~ Features)+
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 3.5,
            position = position_dodge(width = 1),
            vjust = -5.5) +
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "none", axis.title = element_text(size = 20), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13),strip.text.x = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
```


