---
title: "Analysis"
output:  
  html_document:
    self_contained: yes
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)
library(dplyr)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_csv("exp_data.csv")
```


# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


# Data preperation

```{r, echo = FALSE}
# reorder factor 

# to create chart, the data must be in long format and only contain the relevant dependent variables

# 1. make a subset with only the relevant dvs

tdata_long <- tdata


tdata_long$scale_orient <- factor(tdata_long$scale_orient, levels = c("narrow_left", "narrow_right"), 
                               labels = c("narrow_left", "narrow_right"))



# Recode ratings so that high values represent broad scope and low values represent narrow scope

dat_narrow_left <- subset(tdata_long, scale_orient == "narrow_left")
dat_narrow_left$rating_rec <- (dat_narrow_left$DV_rating * 1)

dat_narrow_right <- subset(tdata_long, scale_orient == "narrow_right")
dat_narrow_right$rating_rec <- (1 - dat_narrow_right$DV_rating)

tdata_long <- rbind(dat_narrow_left, dat_narrow_right)


tdata_long$rating_rec <- factor(tdata_long$rating_rec, levels = c(1, 0), labels = c("broad", "narrow"))


library(stringr)
Features <- str_split_fixed(tdata_long$Scenario, "_", 2)[,1]
Knowledge <- str_split_fixed(tdata_long$Scenario, "_", 2)[,2]

tdata_long$Features <- Features
tdata_long$Knowledge <- Knowledge


table(tdata_long$Features, tdata_long$Knowledge)

#write.table(tdata_long, "data_subjects_explans.txt", append = FALSE, sep = "\t", dec = ".",
#            row.names = FALSE, col.names = TRUE)

#write.csv(tdata_long,"tdata_final_explan_analysis.csv", row.names = FALSE)
```
# Analyses

```{r}
# create a summary dataset that also contains the percentages

plotdata <- tdata_long %>%
  group_by(Features, Knowledge, rating_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_sub <- subset(plotdata, rating_rec == "narrow")
```
Tests:

```{r}
(t1 <- binom.test(plotdata$n[1], 40))

(t2 <- binom.test(plotdata$n[2], 40))

(t3 <- binom.test(plotdata$n[3], 40))

(t4 <- binom.test(plotdata$n[4], 40))

(t5 <- binom.test(plotdata$n[5], 40))

(t6 <- binom.test(plotdata$n[6], 40))

(t7 <- binom.test(plotdata$n[7], 40))

(t8 <- binom.test(plotdata$n[8], 40))

(ci_low <- c(t1$conf.int[1], t2$conf.int[1], t3$conf.int[1], t4$conf.int[1], t5$conf.int[1], t6$conf.int[1], t7$conf.int[1], t8$conf.int[1]))

(ci_up <- c(t1$conf.int[2], t2$conf.int[2], t3$conf.int[2], t4$conf.int[2], t5$conf.int[2], t6$conf.int[2], t7$conf.int[2], t8$conf.int[2]))

(p_values <- c(t1$p.value, t2$p.value, t3$p.value, t4$p.value, t5$p.value, t6$p.value, t7$p.value, t8$p.value))

```



```{r}
plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
plotdata$p_value <- p_values

plotdata$ci_low_round <- round(ci_low, 2)
plotdata$ci_up_round <- round(ci_up, 2)
plotdata$p_value <- round(p_values, 6)

selection_tab <- plotdata

selection_tab <- subset(selection_tab, select = c(1:5,9:11))

colnames(selection_tab) <- c('Features','Latent feature info','selection', 'n', 'proportion', '95% CI low', '95% CI up', 'p')



library(kableExtra)

selection_tab %>%
  kbl(caption="Caption",
      format="html",
      col.names = c('Features','Latent feature info','selection', 'n', 'proportion', '95% CI low', '95% CI up', 'p'),
      align="r") %>%
  kable_minimal(full_width = F,  html_font = "Source Sans Pro") -> tab
```


## Graph


```{r, echo = FALSE}
# create a theme to handle the design 
myTheme <- theme(plot.title = element_text(face="bold", size = 16),
        axis.title.x = element_text(face = "bold", size = 20),
        axis.title.y = element_text(face = "bold", size = 20),
        axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 18, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        #panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))


theme_set(theme_light(base_size = 20, base_family = "Poppins"))



library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

#pvalues_x <- c(1:7)
#pvalues <- c(rep("p < .001",6), "p = .057")


g<- ggplot(plotdata, 
       aes(x = Knowledge,
           y = pct,
           fill = rating_rec)) +
  facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1", labels = c("broad scope", "narrow scope")) +
  labs(y = "Percentage", 
       fill = "Selected category",
       x = "Information about latent feature")+ #,
       #title = "Subjects' categorization judgments")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
ggsave("selections.svg",width=6,height=5)
```

# Graphs based on subjects explanations

## What subjects would have done if they hadn't been forced to make a choice


```{r}
library(readr)
explan_data <- read_delim("subjects_explanations_analysis.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)


colnames(explan_data)[3]  <- "equally_likely"               # subjects who wrote explanations saying that 50:50 would have been their preferred answer
colnames(explan_data)[4]  <- "absence_because_visible"      # subjects who inferred the absence of the latent effect because it should have been easily visible
colnames(explan_data)[5]  <- "absence_because_description"  # subjects who inferred the absence of the latent effect given the way the info was presented 
colnames(explan_data)[6]  <- "inferred_absence"             # overall number of subjects who inferred the absence of the latent feature 




explan_data$equally_likely[explan_data$equally_likely == 1] <- "equally_likely"
explan_data$equally_likely[is.na(explan_data$equally_likely)] <- "not_equally_likely_or_unclear"


explan_data$equally_likely <- factor(explan_data$equally_likely, levels = c("not_equally_likely_or_unclear", "equally_likely"), 
                                     labels = c("other", "equally likely"))



explan_data$absence_because_visible[explan_data$absence_because_visible == 1] <- "yes"
explan_data$absence_because_visible[is.na(explan_data$absence_because_visible)] <- "other"


explan_data$absence_because_visible <- factor(explan_data$absence_because_visible, levels = c("other", "yes"), 
                                     labels = c("other", "absence because it should be visible"))



explan_data$absence_because_description[explan_data$absence_because_description == 1] <- "yes"
explan_data$absence_because_description[is.na(explan_data$absence_because_description)] <- "other"


explan_data$absence_because_description <- factor(explan_data$absence_because_description, levels = c("other", "yes"), 
                                     labels = c("other", "absence given the description"))



explan_data$inferred_absence[explan_data$inferred_absence == 1] <- "yes"
explan_data$inferred_absence[is.na(explan_data$inferred_absence)] <- "other"

explan_data$inferred_absence <- factor(explan_data$inferred_absence, levels = c("other", "yes"),
                                       labels = c("other", "inferred absence"))


```



## Subjects clearly stating they'd prefer to answer 50:50


```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, Knowledge, equally_likely) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_sub <- subset(plotdata_between, equally_likely == "equally_likely")


```


```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Knowledge,
           y = pct,
           fill = equally_likely)) +
  facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1", labels = c("other", "'equally likely option'")) +
  labs(y = "Percentage", 
       fill = "Categorization of \nsubjects' explanations",
       x = "Information about latent feature")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
ggsave("preferred_answer.svg",width=6,height=5)
```


## Subjects who inferred the absence of the latent feature

Because of its visibility: 

```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, Knowledge, absence_because_visible) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_sub <- subset(plotdata_between, absence_because_visible == "absence because it should be visible")


```



```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Knowledge,
           y = pct,
           fill = absence_because_visible)) +
  facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Inferred state",
       x = "Information about latent feature",
       title = "Subjects' inference about latent feature")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
ggsave("absence_because_visible.svg",width=6,height=5)
```


Because of how the information was described: 

```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, Knowledge, absence_because_description) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_sub <- subset(plotdata_between, absence_because_description == "absence given the description")


```


```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Knowledge,
           y = pct,
           fill = absence_because_description)) +
  facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Inferred state",
       x = "Information about latent feature",
       title = "Subjects' inference about latent feature 2")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
ggsave("absence_because_description.svg",width=6,height=5)
```

Inferred Absence of latent feature overall: 

```{r}
# create a summary dataset that also contains the percentages

plotdata_between <- explan_data %>%
  group_by(Features, Knowledge, inferred_absence) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata_sub <- subset(plotdata_between, inferred_absence == "inferred absence")


```


```{r}

plotdata <- plotdata_between



g<- ggplot(plotdata, 
       aes(x = Knowledge,
           y = pct,
           fill = inferred_absence)) +
  facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Inferred state",
       x = "Information about latent feature",
       title = "Subjects' inference about latent feature (overall)")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
   #        ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
    #       ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
     #      colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections.pdf",width=6,height=5)
ggsave("absence_overall.svg",width=6,height=5)
```

