---
title: "Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_csv("exp_data.csv")
```


# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


Check n in each condition:

```{r}
table(tdata$condition)
```
There are 8 conditions because what was manipulated whether rate information was given or not, the latent feature (beak vs. feet), varied and counterbalanced the orientation of the rating scale (narrow left vs. narrow right).

```{r}
# make rate information factor a factor (its called "prevalence" in the data frame)

tdata$prevalence <- factor(tdata$prevalence, levels = c("omitted", "presented"), labels = c("omitted", "presented"))

```




# Graphs and summaries

## Rating distribution

```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 14, angle = 0), 
        axis.text.y = element_text(size = 16, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))

tdata_long <- tdata
tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating_rec, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

# new labes for the facets 

g <- ggplot(tdata_sub, aes(x = prevalence, y = valueJitter)) +
  guides(fill=FALSE)+
  #facet_grid( ~ Knowledge)+
  #ggtitle("Subjects' causal srength ratings") +
  scale_y_continuous(limits = c(-5.3, 5.3), breaks=seq(-5, 5, 1), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Spear and \nnet", "Feathers and \ntooth")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black", alpha =0.5) +
  geom_violinhalf(aes(y = rating_rec, group = prevalence, fill = prevalence), color = NA, 
                  position=position_dodge(1), alpha = 0.4)+
  #geom_line(position = pd, color = "black", size = 1, alpha=0.04) +
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  geom_jitter(aes(color = prevalence), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating_rec, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating_rec, group=1), fun.y=mean, geom="line", 
               color = "black", shape = 22, size = 1, alpha = .7)+
  stat_summary(aes(y = rating_rec, group=1, fill = prevalence), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 2, group=1, alpha = 1)+
  stat_summary(aes(y = rating_rec,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Feature rate information", y = "Explanation rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  annotate("text", x = 0.5, y = 3.5, label = c("broad-scope"), angle = 90)+
  annotate("text", x = 0.5, y = -3.5, label = c("narrow-scope"), angle = 90)+
  theme(legend.position = "none")+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)

g

ggsave("results_means.svg",width=4.5,height=5)
ggsave("results_means.pdf",width=4.5,height=5)
#ggsave("results_means_selection.pdf",width=11,height=5)
```

```{r}
library(ggridges)
g2 <- ggplot(tdata_long, aes(x = rating_rec, y = prevalence, fill = prevalence)) +
  scale_x_continuous(breaks = seq(-5, 5, 1))+
  geom_density_ridges(alpha = 0.5)+
   #stat_summary(aes(x = rating_rec), fun.x=mean, geom="point", 
  #             color = "black", shape = 22, size = 2, group=1, alpha = 1)+
  scale_fill_manual(values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_viridis_c(name = "Explanation \nRating", option = "C", breaks=c(-5,0,5), labels=c("narrow scope", "no preference", "broad scope"))+
  labs(x = "Explanation rating", y = "Test question") +
  scale_y_discrete(limits=rev)+
  myTheme+
  theme_light(base_family = "Poppins", base_size = 20)+
  theme(panel.grid = element_blank(), axis.text = element_text(colour ="black"))+
  theme(legend.position="none",
        legend.title=element_blank(),legend.key.width = unit(1.95, 'cm'))+
  annotate("text", y = 0.7, x = 4, label = c("broad-scope"), angle = 0)+
  annotate("text", y = 0.7, x = -4, label = c("narrow-scope"), angle = 0)+
  theme(axis.text.y = element_text(size = 14, angle = 0))

g2

ggsave("results_dist.svg",width=7,height=5)
ggsave("results_dist.pdf",width=6,height=5)
```


# Test difference of group means 

Note that with $N~= 560$, the sample size is at $0.7N$ of the planned final sample size ($N=800$). At this point, we're thus at the planned interim look. 

The critical p-value (one-sided) at this point:

```{r}
library(rpact)
design <- getDesignGroupSequential(
  typeOfDesign = "asP", 
  informationRates = c(0.7, 1), 
  alpha = 0.05, 
  sided = 1)
design$stageLevels * 1
```
is $p~= .0395$. 

Conduct t-Test: 

first, check if variances are equal or not:

```{r}
(tdata_long %>%
  group_by(prevalence) %>%
  summarise_at(vars(rating_rec), list(name=var)) -> var)
```

Variances unequal, so apply correction in T-Test function below:

```{r}
t.test(rating_rec ~ prevalence, data = tdata_long, var.equal = F, alternative = "greater", paired = F) # two sided is used here to get the full confidence interval of the difference; note, however, that the hypothesis was a directed one (thus, alternative = "less" could have been used to get the one-sided p-value). 
```
empirical $p$ is smaller than the critical $p$ determined for the interim look. This also means that data collection can be terminated at this point.

Cohen's d: 

```{r}
# Cohen's d
lsr::cohensD(tdata_long$rating_rec ~ tdata_long$prevalence)
```



Get the values of the CIs of the means:

```{r}
# check width of the means 95% cis (is not supposed to be grater than 0.5)
library(rcompanion)
ci_table <- groupwiseMean(rating_rec ~ prevalence,
              data        = tdata,
              traditional = FALSE,
              percentile  = TRUE)


ci_table

(ci_width <- ci_table$Percentile.upper - ci_table$Percentile.lower)
```

get group medians: 

```{r}
# groupwiseMean(rating_rec ~ Features + Knowledge,
#               data        = tdata_long,
#               traditional = FALSE,
#               percentile  = TRUE)

groupwiseMedian(rating_rec ~ prevalence,
                data        = tdata_long,
                bca         = FALSE,
                percentile  = TRUE,
                R           = 1000)
```


# Subject categories

```{r}
counts <- tdata_long %>%
  group_by(prevalence, rating_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts
# shows that 0 is the mode in all conditions
```



```{r}

tdata_long$category[tdata_long$rating_rec < 0] <- "narrow"
tdata_long$category[tdata_long$rating_rec == 0] <- "unbiased"
tdata_long$category[tdata_long$rating_rec > 0] <- "broad"

counts2 <- tdata_long %>%
  group_by(prevalence, category) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts2

counts2$category <- factor(counts2$category, levels = c("unbiased", "narrow", "broad"), labels = c("unbiased", "narrow l.s.", "broad l.s."))
counts2
```


Get proportion CIs for different conditions:

```{r}
library(PropCIs)
library(DescTools)
library(purrr)

counts_presented <- subset(counts2, prevalence == "presented")

counts_omitted <- subset(counts2, prevalence == "omitted")


(MultinomCI(counts_omitted$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_1)
  

(MultinomCI(counts_presented$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_2)

  




ci_low <- c(selection_ci_1[,2], selection_ci_2[,2])

ci_up <- c(selection_ci_1[,3], selection_ci_2[,3])


```


```{r}
plotdata <- counts2

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

# new labels for the facets 
rates.labs <- c("feature rate information: \nomitted", "feature rate information: \npresented")
names(rates.labs) <- c("omitted", "presented")


g<- ggplot(plotdata, 
       aes(x = category,
           y = pct,
           fill = prevalence)) +
  facet_grid( ~ prevalence, labeller = labeller(prevalence = rates.labs))+
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4,
            position = position_dodge(width = 1),
            vjust = -2.6) +
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "none", axis.title = element_text(size = 20), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13),strip.text.x = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections_between.pdf",width=6,height=5)
ggsave("categories.svg",width=7,height=5)
ggsave("categories.pdf",width=7,height=5)
```

Proportion tests: 

1. Test if proportion of normative judgments is higher in "omitted"

```{r}
normative <- prop.test(x = c(plotdata$n[3], plotdata$n[6]), n = c(280, 280), alternative = "greater", correct = F)
normative
```
2. Test if proportion of narrow latent scope biases is higher in "presented"

```{r}
narrowlsb <- prop.test(x = c(plotdata$n[5], plotdata$n[2]), n = c(280, 280), alternative = "greater", correct = F)
narrowlsb
```


# Remaining results (additional test cases)

Blue feet probe:

```{r}
counts_bluefeet <- tdata_long %>%
  group_by(bluefeet_choice) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts_bluefeet

counts_bluefeet$bluefeet_choice <- factor(counts_bluefeet$bluefeet_choice, 
                                          levels = c("This duck has no mutation affecting beak and feet color.",
                                                     "This duck has a TOX20 mutation (blue beak).",
                                                     "This duck has a NAX20 mutation (blue feet).",
                                                     "This duck has a PIX67 mutation (blue beak and blue feet).",
                                                     "This duck has a TOX20 mutation (blue beak) and a NAX20 mutation (blue feet).",
                                                     "This duck has PIX67 mutation (blue beak and blue feet) and a TOX20 mutation (blue beak).",
                                                     "This duck has a PIX67 mutation (blue beak and blue feet) and NAX20 mutation (blue feet)."
                                                     ),
                                          labels = c("no mutation", "mutation affecting \nonly beak", "mutation affecting \nonly feet",
                                                     "mutation affecting \nbeak and feet", 
                                                     "mutation affecting \nonly beak + \nmutation affecting \nonly feet", 
                                                     "mutation affecting \nbeak and feet + \nmutation affecting \nonly beak",
                                                     "mutation affecting \nbeak and feet + \nmutation affecting \nonly feet"))


```


```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts_bluefeet$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts_bluefeet

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = bluefeet_choice,
           y = pct,
           fill = bluefeet_choice)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels=c("BkFt2", "Ft1")) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4,
            position = position_dodge(width = 1),
            vjust = -1.5) +
  scale_fill_manual(name = "Selected explanation",values=c("#66c2a5", "#e78ac3", "#e6ab02", "#8da0cb", "#a6d854", "#a65628", "#e41a1c"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "right", axis.title = element_blank(), axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 12),legend.title = element_text(size = 13), legend.spacing.y = unit(0.3, 'cm'))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())+
    guides(fill = guide_legend(byrow = TRUE))

g

#ggsave("selections_between.pdf",width=6,height=5)
ggsave("categories_bluefeet.svg",width=8,height=5)
ggsave("categories_bluefeet.pdf",width=8,height=5)
```





Blue beak probe: 


```{r}
counts_bluebeak <- tdata_long %>%
  group_by(bluebeak_choice) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct, accuracy = 0.1))

counts_bluebeak



# add the missing categories to the data frame (to have them in the plot later)

counts_bluebeak[nrow(counts_bluebeak) + 1,] <- list("This duck has a PIX67 mutation (blue beak and blue feet) and NAX20 mutation (blue feet).", 0, 0.00, "0%")


counts_bluebeak[nrow(counts_bluebeak) + 1,] <- list("This duck has PIX67 mutation (blue beak and blue feet) and a TOX20 mutation (blue beak).", 0, 0.00, "0%")


counts_bluebeak


counts_bluebeak$bluebeak_choice <- factor(counts_bluebeak$bluebeak_choice, 
                                          levels = c("This duck has no mutation affecting beak and feet color.",
                                                     "This duck has a TOX20 mutation (blue beak).",
                                                     "This duck has a NAX20 mutation (blue feet).",
                                                     "This duck has a PIX67 mutation (blue beak and blue feet).",
                                                     "This duck has a TOX20 mutation (blue beak) and a NAX20 mutation (blue feet).",
                                                     "This duck has PIX67 mutation (blue beak and blue feet) and a TOX20 mutation (blue beak).",
                                                     "This duck has a PIX67 mutation (blue beak and blue feet) and NAX20 mutation (blue feet)."
                                                     ),
                                          labels = c("no mutation", "mutation affecting \nonly beak", "mutation affecting \nonly feet",
                                                     "mutation affecting \nbeak and feet", 
                                                     "mutation affecting \nonly beak + \nmutation affecting \nonly feet", 
                                                     "mutation affecting \nbeak and feet + \nmutation affecting \nonly beak",
                                                     "mutation affecting \nbeak and feet + \nmutation affecting \nonly feet"))


```


```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts_bluebeak$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts_bluebeak

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = bluebeak_choice,
           y = pct,
           fill = bluebeak_choice)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels=c("Bk1", "Ft1")) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4,
            position = position_dodge(width = 1),
            vjust = -1.5) +
  scale_fill_manual(name = "Selected explanation",values=c("#66c2a5", "#e78ac3", "#e6ab02", "#8da0cb", "#a6d854", "#a65628", "#e41a1c"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "right", axis.title = element_blank(), axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 12),legend.title = element_text(size = 13), legend.spacing.y = unit(0.3, 'cm'))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())+
    guides(fill = guide_legend(byrow = TRUE))

g

#ggsave("selections_between.pdf",width=6,height=5)
ggsave("categories_bluebeak.svg",width=8,height=5)
ggsave("categories_bluebeak.pdf",width=8,height=5)
```






typical duck probe: 


```{r}
counts_typical <- tdata_long %>%
  group_by(typical_choice) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts_typical


# add the missing categories to the data frame (to have them in the plot later)


counts_typical[nrow(counts_typical) + 1,] <- list("This duck has a PIX67 mutation (blue beak and blue feet) and NAX20 mutation (blue feet).", 0, 0.00, "0%")

counts_typical[nrow(counts_typical) + 1,] <- list("This duck has a TOX20 mutation (blue beak).", 0, 0.00, "0%")


counts_typical


counts_typical$typical_choice <- factor(counts_typical$typical_choice, 
                                          levels = c("This duck has no mutation affecting beak and feet color.",
                                                     "This duck has a TOX20 mutation (blue beak).",
                                                     "This duck has a NAX20 mutation (blue feet).",
                                                     "This duck has a PIX67 mutation (blue beak and blue feet).",
                                                     "This duck has a TOX20 mutation (blue beak) and a NAX20 mutation (blue feet).",
                                                     "This duck has PIX67 mutation (blue beak and blue feet) and a TOX20 mutation (blue beak).",
                                                     "This duck has a PIX67 mutation (blue beak and blue feet) and NAX20 mutation (blue feet)."
                                                     ),
                                          labels = c("no mutation", "mutation affecting \nonly beak", "mutation affecting \nonly feet",
                                                     "mutation affecting \nbeak and feet", 
                                                     "mutation affecting \nonly beak + \nmutation affecting \nonly feet", 
                                                     "mutation affecting \nbeak and feet + \nmutation affecting \nonly beak",
                                                     "mutation affecting \nbeak and feet + \nmutation affecting \nonly feet"))


counts_typical
```


```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts_typical$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts_typical

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = typical_choice,
           y = pct,
           fill = typical_choice)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels=c("No mutation")) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4,
            position = position_dodge(width = 1),
            vjust = -1.5) +
  scale_fill_manual(name = "Selected explanation",values=c("#66c2a5", "#e78ac3", "#e6ab02", "#8da0cb", "#a6d854", "#a65628", "#e41a1c"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "right", axis.title = element_blank(), axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 12),legend.title = element_text(size = 13), legend.spacing.y = unit(0.3, 'cm'))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())+
    guides(fill = guide_legend(byrow = TRUE))

g

#ggsave("selections_between.pdf",width=6,height=5)
ggsave("categories_typical.svg",width=8,height=5)
ggsave("categories_typical.pdf",width=8,height=5)
```






```{r}
counts_both <- tdata_long %>%
  group_by(bluefeet_bluebeak_choice) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts_both 


counts_both$bluefeet_bluebeak_choice <- factor(counts_both$bluefeet_bluebeak_choice, 
                                          levels = c("This duck has no mutation affecting beak and feet color.",
                                                     "This duck has a TOX20 mutation (blue beak).",
                                                     "This duck has a NAX20 mutation (blue feet).",
                                                     "This duck has a PIX67 mutation (blue beak and blue feet).",
                                                     "This duck has a TOX20 mutation (blue beak) and a NAX20 mutation (blue feet).",
                                                     "This duck has PIX67 mutation (blue beak and blue feet) and a TOX20 mutation (blue beak).",
                                                     "This duck has a PIX67 mutation (blue beak and blue feet) and NAX20 mutation (blue feet)."
                                                     ),
                                          labels = c("no mutation", "mutation affecting \nonly beak", "mutation affecting \nonly feet",
                                                     "mutation affecting \nbeak and feet", 
                                                     "mutation affecting \nonly beak + \nmutation affecting \nonly feet", 
                                                     "mutation affecting \nbeak and feet + \nmutation affecting \nonly beak",
                                                     "mutation affecting \nbeak and feet + \nmutation affecting \nonly feet"))


counts_both


```


```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts_both$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts_both

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = bluefeet_bluebeak_choice,
           y = pct,
           fill = bluefeet_bluebeak_choice)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels=c("BkFt2", "Bk1 + Ft1", "Bk1", "BkFt2 + Ft1", "BkFt2")) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4,
            position = position_dodge(width = 1),
            vjust = -1.5) +
  scale_fill_manual(name = "Selected explanation",values=c("#66c2a5", "#e78ac3", "#e6ab02", "#8da0cb", "#a6d854", "#a65628", "#e41a1c"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "right", axis.title = element_blank(), axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 12),legend.title = element_text(size = 13), legend.spacing.y = unit(0.3, 'cm'))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())+
    guides(fill = guide_legend(byrow = TRUE))

g

#ggsave("selections_between.pdf",width=6,height=5)
ggsave("categories_both.svg",width=8,height=5)
ggsave("categories_both.pdf",width=8,height=5)
```


