---
title: "Explanation Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)
library(scales)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
explan_data <- read_delim("subject_explanations.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
```



# Data preparation

```{r, echo = FALSE}
# reorder factor 

# to create chart, the data must be in long format and only contain the relevant dependent variables

# 1. make a subset with only the relevant dvs

tdata_long <- explan_data


tdata_long$scale_orient <- factor(tdata_long$scale_orient, levels = c("narrow_left", "narrow_right"), 
                               labels = c("narrow_left", "narrow_right"))




tdata_long$`Correct explanation`[is.na(tdata_long$`Correct explanation`)] <- "0"
tdata_long$`Unclear explanation`[is.na(tdata_long$`Unclear explanation`)] <- "0"
tdata_long$`Inferred absence because of low feature base rate`[is.na(tdata_long$`Inferred absence because of low feature base rate`)] <- "0"
tdata_long$`Inferred absence of latent feature`[is.na(tdata_long$`Inferred absence of latent feature`)] <- "0"


```


Note: Subjects' explanations were coded in the last columns (the column names describe the coded criterion). 1 means that the coding criterion is (clearly) met, 0 that it isn't. 

```{r}
# turn explanation coding values into factors 

explan_data <- tdata_long

explan_data$`Correct explanation` <- as.factor(explan_data$`Correct explanation`)

explan_data$`Unclear explanation` <- as.factor(explan_data$`Unclear explanation`)

explan_data$`Inferred absence because of low feature base rate` <- as.factor(explan_data$`Inferred absence because of low feature base rate`)

explan_data$`Inferred absence of latent feature` <- as.factor(explan_data$`Inferred absence of latent feature`)

```


# Explanation Analysis

1. Proportions of subjects who gave (correct) explanations for why the answer is 50:50

```{r}
# create a summary dataset that also contains the percentages
plotdata_between <- explan_data %>%
  group_by(prevalence, `Correct explanation`) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))


plotdata_between

plotdata_sub <- subset(plotdata_between, `Correct explanation` == 1)
```



```{r}
plotdata <- plotdata_between

g<- ggplot(plotdata, 
       aes(x = "",
           y = pct,
           fill = `Correct explanation`)) +
  facet_grid( ~ prevalence)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Correct Explanation",
       x = "Condition")+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
```

```{r}
normative <- prop.test(x = c(plotdata_between$n[2], plotdata_between$n[4]), n = c(280, 280), alternative = "two.sided", correct = F)
normative
```

2. proportions of subjects who inferred the absence of the unobserved feature due to its low overall probability (inferred evidence account)

```{r}
# create a summary dataset that also contains the percentages
plotdata_between <- explan_data %>%
  group_by(prevalence, `Inferred absence because of low feature base rate`) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))


plotdata_between

plotdata_sub <- subset(plotdata_between, `Inferred absence because of low feature base rate` == 1)
```


```{r}
plotdata <- plotdata_between

g<- ggplot(plotdata, 
       aes(x = "",
           y = pct,
           fill = `Inferred absence because of low feature base rate`)) +
  facet_grid( ~ prevalence)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Correct Explanation",
       x = "Features")+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
```

```{r}
rate <- prop.test(x = c(0, plotdata_between$n[3]), n = c(280, 280), alternative = "less", correct = F)
rate
```
