---
title: "Sample Size Planning"
output: html_document
---

# Calculating a plausible effect size

1. Get an effect size that should be detected reliably (based on expected mean differences and what we know about the SD in this kind of study)


```{r}
library(pwr)
pwr.t.test(d=(-0.2-(-0.5))/1.6,
           power=0.8,
           sig.level=0.05,
           type="two.sample",
           alternative="greater")
```

This yields $N~= 712$ ($n~= 356$), 89 in each of the 8 total conditions. This is the sample size if there was only one look at the data. 

A small effect of $d~= 0.19$ can be detected with 80% power in a one-sided (directed) two-sample t test if the sample size is $N~= 712$.

Note, however, that this sample is to small if a sequential testing strategy is applied. This is because of the Type-I error control in such an analysis, which leads to a reduction in test power. The final necessary total sample size will thus be conducted below.


# Planning with Sequential Testing

However, the present study aimed to employ a sequential testing strategy with one planned interim look after __70%__ of the final sample size.

For more information, see [this website](https://lakens.github.io/statistical_inferences/10-sequential.html). 

What are the critical p-values in this sequential analysis?

```{r}
library(rpact)
design <- getDesignGroupSequential(
  typeOfDesign = "asP", 
  informationRates = c(0.7, 1), 
  alpha = 0.05, 
  sided = 1)
design$stageLevels * 1
```

The critical (on-sided) p-value at the interim look is $.039$, and the critical p-value at the final look is $.0267$. 

Since these p-value adjustments lead to p-values smaller than $0.05$ (the sig. value used in the initial power planning above), this leads to a (mild) reduction of statistical power. This can be compensated, however, by increasing the planned sample size accordingly. 

This can be done this way:

```{r}
library(rpact)
seq_design <- getDesignGroupSequential(
  informationRates = c(0.7, 1), 
  typeOfDesign = "asP",
  sided = 1,
  alpha = 0.05,
  beta = 0.2
  )

# Compute the sample size we need
power_res_seq <- getSampleSizeMeans(
  design = seq_design,
  groups = 2,
  alternative = 0.1875, 
  stDev = 1, 
  allocationRatioPlanned = 1,
  normalApproximation = FALSE)

power_res_seq$maxNumberOfSubjects
```
To compensate for the power loss due to sequential testing with one interim look at $0.7N$, a total sample size of $N~= 789$ is needed. This was rounded to $N~= 800$ ($n~= 100$ per condition). 

## Determine the N for the interim look

Given this information, at what point (i.e., at what fraction of the total N of 800) must the interim look take place? 

```{r}
interim_N <- 0.7 * 800
interim_N
```
Interim look takes place at $N~= 560$ ($n~= 70$ subjects in each of the 8 conditions)


