---
title: "Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_csv("exp_data.csv")
```



```{r}
# remove subject with duplicate entry
tdata <- tdata %>% distinct(subj_code, .keep_all = TRUE)

# Note that two rows appear twice, and the duplicates must be deleted
```



# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


Check n in each condition:

```{r}
table(tdata$condition)
```
There are 4 conditions because we manipulated the latent feature (beak vs. feet) and counterbalanced the orientation of the rating scale (narrow left vs. narrow right)


# Check mean and CI to estimate the bias

```{r}
# check width of the means 95% cis (is not supposed to be grater than 0.5)
library(rcompanion)
ci_table <- groupwiseMean(rating_rec ~ 1,
              data        = tdata,
              traditional = FALSE,
              percentile  = TRUE)


ci_table

(ci_width <- ci_table$Percentile.upper - ci_table$Percentile.lower)
```

# Graphs and summaries

## Rating distribution

```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 14, angle = 0), 
        axis.text.y = element_text(size = 16, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        #panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))

tdata_long <- tdata
tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating_rec, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

# new labes for the facets 

g <- ggplot(tdata_sub, aes(x = 0.5, y = valueJitter)) +
  guides(fill=FALSE)+
  #facet_grid( ~ dv_query)+
  #ggtitle("Subjects' causal srength ratings") +
  scale_y_continuous(limits = c(-5.3, 5.3), breaks=seq(-5, 5, 1), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Probe")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black", alpha =0.5) +
  geom_violinhalf(aes(y = rating_rec, group = NA, fill = "#66c2a5"), color = NA, 
                  position=position_dodge(1), alpha = 0.4)+
  #geom_line(position = pd, color = "black", size = 1, alpha=0.04) +
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  geom_jitter(aes(color = "#66c2a5"), alpha = 0.5, width = 0.15, height = 0.3) +
  stat_summary(aes(y = rating_rec, group=NA), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating_rec, group=1, fill = "#66c2a5"), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 2, group=1, alpha = 1)+
  stat_summary(aes(y = rating_rec,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "", y = "Explanation rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  annotate("text", x = 0, y = 3.5, label = c("broad-scope"), angle = 90)+
  annotate("text", x = 0, y = -3.5, label = c("narrow-scope"), angle = 90)+
  theme(legend.position = "none")+
  myTheme+
   theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'),
         axis.text.x = element_blank(),
         axis.title.x = element_blank(),
         axis.ticks.x = element_blank())+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -8)
g


ggsave("results_means_mainDV.svg",width=5,height=5)
ggsave("results_means_mainDV.pdf",width=5,height=5)
```


## Subject categories

Get the values of the CIs shown in the plot:

```{r}
values <- ggplot_build(g)$data[[4]] # values are shown in the 4th panel 
values
```


get group medians: 

```{r}
# groupwiseMean(rating_rec ~ Features + Knowledge,
#               data        = tdata_long,
#               traditional = FALSE,
#               percentile  = TRUE)

groupwiseMedian(rating_rec ~ 1,
                data        = tdata_long,
                bca         = FALSE,
                percentile  = TRUE,
                R           = 1000)
```




```{r}
counts <- tdata_long %>%
  group_by(rating_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts
# shows that 0 is the mode in all conditions
```



```{r}

tdata_long$category[tdata_long$rating_rec < 0] <- "narrow"
tdata_long$category[tdata_long$rating_rec == 0] <- "unbiased"
tdata_long$category[tdata_long$rating_rec > 0] <- "broad"

counts2 <- tdata_long %>%
  group_by(category) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts2

counts2$category <- factor(counts2$category, levels = c("unbiased", "narrow", "broad"), labels = c("unbiased", "narrow l.s.", "broad l.s."))

```


Get proportion CIs for the 3 categories

```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts2$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts2

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = category,
           y = pct,
           fill = "#66c2a5")) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 3.5,
            position = position_dodge(width = 1),
            vjust = -3) +
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "none", axis.title = element_text(size = 20), axis.text = element_text(size = 11, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g


ggsave("categories.svg",width=5,height=5)
ggsave("categories.pdf",width=5,height=5)
```

# Remaining results (additional test cases)

Blue feet probe:

```{r}
counts_bluefeet <- tdata_long %>%
  group_by(bluefeet_choice) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts_bluefeet

# add the missing categories to the data frame (to have them in the plot later)

#This duck has a Bk1 mutation (blue beak).
counts_bluefeet[nrow(counts_bluefeet) + 1,] <- list("Bk1", 0, 0.00, "0%")

# This duck has no mutation affecting beak and feet color.
counts_bluefeet[nrow(counts_bluefeet) + 1,] <- list("no mutation", 0, 0.00, "0%")

#This duck has a BkFt2 mutation (blue beak and blue feet) and Ft1 mutation (blue feet).
counts_bluefeet[nrow(counts_bluefeet) + 1,] <- list("BkFt2 + Ft1", 0, 0.00, "0%")

#This duck has BkFt2 mutation (blue beak and blue feet) and a Bk1 mutation (blue beak).
counts_bluefeet[nrow(counts_bluefeet) + 1,] <- list("BkFt2 + Bk1", 0, 0.00, "0%")

#This duck has a Bk1 mutation (blue beak only) and a Ft1 mutation (blue feet only).
counts_bluefeet[nrow(counts_bluefeet) + 1,] <- list("Bk1 + Ft1", 0, 0.00, "0%")



counts_bluefeet


counts_bluefeet$bluefeet_choice <- factor(counts_bluefeet$bluefeet_choice, 
                                          levels = c("no mutation",
                                                     "Bk1",
                                                     "This duck has a Ft1 mutation (blue feet).",
                                                     "This duck has a BkFt2 mutation (blue beak and blue feet).",
                                                     "Bk1 + Ft1",
                                                     "BkFt2 + Bk1",
                                                     "BkFt2 + Ft1"
                                                     ),
                                          labels = c("no mutation", "Bk1", "Ft1", "BkFt2", "Bk1 + Ft1", "BkFt2 + Bk1", "BkFt2 + Ft1"))


```


```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts_bluefeet$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts_bluefeet

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = bluefeet_choice,
           y = pct,
           fill = bluefeet_choice)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels=c("BkFt2", "Ft1")) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 5,
            position = position_dodge(width = 1),
            vjust = -1.5) +
  scale_fill_manual(name = "Selected explan.",values=c("#66c2a5", "#e78ac3", "#e6ab02", "#8da0cb", "#a6d854", "#a65628", "#e41a1c"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "right", axis.title = element_blank(), axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())

g


ggsave("categories_bluefeet.svg",width=7,height=4)
ggsave("categories_bluefeet.pdf",width=7,height=4)
```





Blue beak probe: 


```{r}
counts_bluebeak <- tdata_long %>%
  group_by(bluebeak_choice) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct, accuracy = 0.1))

counts_bluebeak



# add the missing categories to the data frame (to have them in the plot later)


counts_bluebeak[nrow(counts_bluebeak) + 1,] <- list("BkFt2", 0, 0.00, "0%")


counts_bluebeak[nrow(counts_bluebeak) + 1,] <- list("no mutation", 0, 0.00, "0%")


counts_bluebeak[nrow(counts_bluebeak) + 1,] <- list("BkFt2 + Ft1", 0, 0.00, "0%")


counts_bluebeak[nrow(counts_bluebeak) + 1,] <- list("BkFt2 + Bk1", 0, 0.00, "0%")


counts_bluebeak[nrow(counts_bluebeak) + 1,] <- list("Bk1 + Ft1", 0, 0.00, "0%")



counts_bluebeak


counts_bluebeak$bluebeak_choice <- factor(counts_bluebeak$bluebeak_choice, 
                                          levels = c("no mutation",
                                                     "This duck has a Bk1 mutation (blue beak).",
                                                     "This duck has a Ft1 mutation (blue feet).",
                                                     "BkFt2",
                                                     "Bk1 + Ft1",
                                                     "BkFt2 + Bk1",
                                                     "BkFt2 + Ft1"
                                                     ),
                                          labels = c("no mutation", "Bk1", "Ft1", "BkFt2", "Bk1 + Ft1", "BkFt2 + Bk1", "BkFt2 + Ft1"))


```


```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts_bluebeak$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts_bluebeak

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = bluebeak_choice,
           y = pct,
           fill = bluebeak_choice)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels=c("Bk1", "Ft1")) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 5,
            position = position_dodge(width = 1),
            vjust = -1.5) +
  scale_fill_manual(name = "Selected explan.",values=c("#66c2a5", "#e78ac3", "#e6ab02", "#8da0cb", "#a6d854", "#a65628", "#e41a1c"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "right", axis.title = element_blank(), axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())

g

ggsave("categories_bluebeak.svg",width=7,height=4)
ggsave("categories_bluebeak.pdf",width=7,height=4)
```






typical duck probe: 


```{r}
counts_typical <- tdata_long %>%
  group_by(typical_choice) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts_typical


# add the missing categories to the data frame (to have them in the plot later)


counts_typical[nrow(counts_typical) + 1,] <- list("BkFt2", 0, 0.00, "0%")


counts_typical[nrow(counts_typical) + 1,] <- list("Bk1", 0, 0.00, "0%")


counts_typical[nrow(counts_typical) + 1,] <- list("Ft1", 0, 0.00, "0%")


counts_typical[nrow(counts_typical) + 1,] <- list("BkFt2 + Ft1", 0, 0.00, "0%")


counts_typical[nrow(counts_typical) + 1,] <- list("BkFt2 + Bk1", 0, 0.00, "0%")


counts_typical[nrow(counts_typical) + 1,] <- list("Bk1 + Ft1", 0, 0.00, "0%")



counts_typical


counts_typical$typical_choice <- factor(counts_typical$typical_choice, 
                                          levels = c("This duck has no mutation affecting beak and feet color.",
                                                     "Bk1",
                                                     "Ft1",
                                                     "BkFt2",
                                                     "Bk1 + Ft1",
                                                     "BkFt2 + Bk1",
                                                     "BkFt2 + Ft1"
                                                     ),
                                          labels = c("no mutation", "Bk1", "Ft1", "BkFt2", "Bk1 + Ft1", "BkFt2 + Bk1", "BkFt2 + Ft1"))


counts_typical
```


```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts_typical$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts_typical

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = typical_choice,
           y = pct,
           fill = typical_choice)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels=c("No mutation")) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 5,
            position = position_dodge(width = 1),
            vjust = -1.5) +
  scale_fill_manual(name = "Selected explan.",values=c("#66c2a5", "#e78ac3", "#e6ab02", "#8da0cb", "#a6d854", "#a65628", "#e41a1c"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "right", axis.title = element_blank(), axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())

g


ggsave("categories_typical.svg",width=7,height=4)
ggsave("categories_typical.pdf",width=7,height=4)
```






```{r}
counts_both <- tdata_long %>%
  group_by(bluefeet_bluebeak_choice) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts_both 

counts_both[nrow(counts_both) + 1,] <- list("Ft1", 0, 0.00, "0%")


counts_both[nrow(counts_both) + 1,] <- list("no mutation", 0, 0.00, "0%")

counts_both


counts_both$bluefeet_bluebeak_choice <- factor(counts_both$bluefeet_bluebeak_choice, 
                                          levels = c("no mutation",
                                                     "This duck has a Bk1 mutation (blue beak).",
                                                     "Ft1",
                                                     "This duck has a BkFt2 mutation (blue beak and blue feet).",
                                                     "This duck has a Bk1 mutation (blue beak only) and a Ft1 mutation (blue feet only).",
                                                     "This duck has BkFt2 mutation (blue beak and blue feet) and a Bk1 mutation (blue beak).",
                                                     "This duck has a BkFt2 mutation (blue beak and blue feet) and Ft1 mutation (blue feet)."
                                                     ),
                                          labels = c("no mutation", "Bk1", "Ft1", "BkFt2", "Bk1 + Ft1", "BkFt2 + Bk1", "BkFt2 + Ft1"))


counts_both


```


```{r}
library(PropCIs)
library(DescTools)
library(purrr)

(MultinomCI(counts_both$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci)
  

ci_low <- c(selection_ci[,2])

ci_up <- c(selection_ci[,3])
```


```{r}
plotdata <- counts_both

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = bluefeet_bluebeak_choice,
           y = pct,
           fill = bluefeet_bluebeak_choice)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels=c("BkFt2", "Bk1 + Ft1", "Bk1", "BkFt2 + Ft1", "BkFt2")) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 5,
            position = position_dodge(width = 1),
            vjust = -1.5) +
  scale_fill_manual(name = "Selected explan.",values=c("#66c2a5", "#e78ac3", "#e6ab02", "#8da0cb", "#a6d854", "#a65628", "#e41a1c"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "right", axis.title = element_blank(), axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())

g

ggsave("categories_both.svg",width=7,height=4)
ggsave("categories_both.pdf",width=7,height=4)
```


