---
title: "Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_csv("exp_data.csv")

```


# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say






# Results

```{r, echo = FALSE}
# reorder factor 

# to create chart, the data must be in long format and only contain the relevant dependent variables

# 1. make a subset with only the relevant dvs

tdata_long <- tdata


tdata_long$scale_orient <- factor(tdata_long$scale_orient, levels = c("narrow_left", "narrow_right"), 
                               labels = c("narrow_left", "narrow_right"))



# Recode ratings so that high values represent broad scope and low values represent narrow scope

dat_narrow_left <- subset(tdata_long, scale_orient == "narrow_left")
dat_narrow_left$rating_rec <- (dat_narrow_left$DV_rating) - 5

dat_narrow_right <- subset(tdata_long, scale_orient == "narrow_right")
dat_narrow_right$rating_rec <- (dat_narrow_right$DV_rating * -1) + 5

tdata_long <- rbind(dat_narrow_left, dat_narrow_right)


library(stringr)
Features <- str_split_fixed(tdata_long$Scenario, "_", 2)[,1]
Knowledge <- str_split_fixed(tdata_long$Scenario, "_", 2)[,2]

tdata_long$Features <- Features
tdata_long$Knowledge <- Knowledge


table(tdata_long$Features, tdata_long$Knowledge)

tdata_long$Features <- factor(tdata_long$Features, levels = c("spearNet", "featherTooth"), labels = c("Spear and net", "Feathers and tooth"))

tdata_long$Knowledge <- factor(tdata_long$Knowledge, levels = c("notKnow", "noInfo"), labels = c("'you don't know'", "no information"))

```

## Graphs

```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 14, angle = 0), 
        axis.text.y = element_text(size = 16, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))

tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating_rec, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

# new labes for the facets 

g <- ggplot(tdata_sub, aes(x = Features, y = valueJitter)) +
  guides(fill=FALSE)+
  facet_grid( ~ Knowledge)+
  #ggtitle("Subjects' causal srength ratings") +
  scale_y_continuous(limits = c(-5.3, 5.3), breaks=seq(-5, 5, 1), expand = c(0,0)) +
  scale_x_discrete(labels=c("Spear and \nnet", "Feather and \ntooth")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black", alpha =0.5) +
  geom_violinhalf(aes(y = rating_rec, group = Features, fill = Features), color = NA, 
                  position=position_dodge(1), alpha = 0.4)+
  #geom_line(position = pd, color = "black", size = 1, alpha=0.04) +
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  geom_jitter(aes(color = Features), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating_rec, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating_rec, group=1, fill = Features), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 2, group=1, alpha = 1)+
  stat_summary(aes(y = rating_rec,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Category Features", y = "Categorization Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  annotate("text", x = 0.5, y = 3.5, label = c("broad-scope"), angle = 90)+
  annotate("text", x = 0.5, y = -3.5, label = c("narrow-scope"), angle = 90)+
  theme(legend.position = "none")+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)

g

ggsave("results_means_mainDV.svg",width=7,height=5)
ggsave("results_means_mainDV.pdf",width=7,height=5)
#ggsave("results_means_selection.pdf",width=11,height=5)
```

Use the ggplot_build package to see a table with the means and CI values plotted in the graph: 

```{r}
values <- ggplot_build(g)$data[[4]] # values are shown in the 4th panel 
values
```

get group medians: 

```{r}
library(rcompanion)
# groupwiseMean(rating_rec ~ Features + Knowledge,
#               data        = tdata_long,
#               traditional = FALSE,
#               percentile  = TRUE)

groupwiseMedian(rating_rec ~ Features + Knowledge,
                data        = tdata_long,
                bca         = FALSE,
                percentile  = TRUE,
                R           = 1000)
```



```{r}
counts <- tdata_long %>%
  group_by(Features, Knowledge, rating_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts
```


```{r}
counts <- tdata_long %>%
  group_by(Features, Knowledge, rating_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts


```

```{r}

tdata_long$category[tdata_long$rating_rec < 0] <- "narrow"
tdata_long$category[tdata_long$rating_rec == 0] <- "unbiased"
tdata_long$category[tdata_long$rating_rec > 0] <- "broad"

counts2 <- tdata_long %>%
  group_by(Features, Knowledge, category) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

counts2

counts2$category <- factor(counts2$category, levels = c("unbiased", "narrow", "broad"), labels = c("unbiased", "narrow l.s.", "broad l.s."))

```


Get proportion CIs for different categories in each diagnosability condition:

```{r}
library(PropCIs)
library(DescTools)
library(purrr)

counts_spear_notknow <- subset(counts2, Features == "Spear and net" & Knowledge == "'you don't know'")
counts_spear_noinf <- subset(counts2, Features == "Spear and net" & Knowledge == "no information")

counts_feather_notknow <- subset(counts2, Features == "Feathers and tooth" & Knowledge == "'you don't know'")
counts_feather_noinf <- subset(counts2, Features == "Feathers and tooth" & Knowledge == "no information")


(MultinomCI(counts_spear_notknow$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_1)
  

(MultinomCI(counts_spear_noinf$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_2)
  
(MultinomCI(counts_feather_notknow$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_3)
  

(MultinomCI(counts_feather_noinf$n,
           conf.level=0.95,
           method="sisonglaz") -> selection_ci_4)
  



ci_low <- c(selection_ci_1[,2], selection_ci_2[,2], selection_ci_3[,2], selection_ci_4[,2])

ci_up <- c(selection_ci_1[,3], selection_ci_2[,3], selection_ci_3[,3], selection_ci_4[,3])


```


```{r}
plotdata <- counts2

plotdata$ci_low <- ci_low
plotdata$ci_up <- ci_up
```


Plot:

```{r}

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = category,
           y = pct,
           fill = category)) +
  facet_grid(Features ~ Knowledge)+
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 3.5,
            position = position_dodge(width = 1),
            vjust = -3.5) +
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Explanatory preference",
       x = "Explanatory preference")+
  geom_pointrange(ymin = ci_low, ymax = ci_up, position = position_dodge(width = 0.89), shape = 22, size = 0.3)+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = plotdata$Transformation, y = plotdata$pct, 
   #        ymin = plotdata$ci_low, 
    #       ymax = plotdata$ci_up, 
     #      colour = "black", size = 0.8, shape = 22, fill = Transformation, fatten = 1)+
  #annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "none", axis.title = element_text(size = 20), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13),strip.text.x = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections_between.pdf",width=6,height=5)
ggsave("categories.svg",width=7,height=5)
ggsave("categories.pdf",width=7,height=5)
```



## Analyses

```{r, echo = FALSE, warning = FALSE, message = FALSE}
################################################################################################################
##################################### Statistical Analyses #####################################################
################################################################################################################
library(pastecs)
library(lme4)
library(nlme)
library(ez)
library(vtable)
library(rcompanion)
```


### Model comparisons


```{r}
library(afex)
library(emmeans)

a1 <- aov_car(rating_rec ~ Features*Knowledge + Error(subj_code), tdata_long, anova_table = list(es = "pes"))
a1

```
more digits for the p.values can be obtained by running model comparisons manually.

```{r}
null <- lme(rating_rec ~ 1, random = ~1|subj_code, data = tdata_long, method = "ML")

main_features <- lme(rating_rec ~ Features, random = ~1|subj_code, data = tdata_long, method = "ML")

anova(null, main_features)
```


```{r}
main_knowledge <- lme(rating_rec ~ Features + Knowledge, random = ~1|subj_code, data = tdata_long, method = "ML")

anova(null, main_features, main_knowledge)
```


```{r}
interaction <- lme(rating_rec ~ Features*Knowledge, random = ~1|subj_code, data = tdata_long, method = "ML")

anova(null, main_features, main_knowledge, interaction)
```



### Contrasts

```{r}
############### 
# a follow-up analysis 
library(lsmeans)
# means

ls2 <- lsmeans(a1, c("Features", "Knowledge")) # group means by between-condition
ls2

# contrast the strength levels (main effect; averaging over decision level, as there was no sig. interaction)
contrasts <- emmeans(a1, ~ Features*Knowledge)
s <- pairs(contrasts, adjust = "none")


s
confint(s, level = 0.95)
```





