---
title: "Explanation Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)
library(scales)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
explan_data <- read_delim("subject_explanations.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
```


# Data preparation

```{r, echo = FALSE}
# reorder factor 

# to create chart, the data must be in long format and only contain the relevant dependent variables

# 1. make a subset with only the relevant dvs

tdata_long <- explan_data


tdata_long$scale_orient <- factor(tdata_long$scale_orient, levels = c("narrow_left", "narrow_right"), 
                               labels = c("narrow_left", "narrow_right"))




tdata_long$dv_query <- factor(tdata_long$dv_query, levels = c("probability", "satisfaction_would", "satisfaction_is"))

tdata_long$`Correct explanation`[is.na(tdata_long$`Correct explanation`)] <- "0"
tdata_long$`Unclear explanation`[is.na(tdata_long$`Unclear explanation`)] <- "0"
tdata_long$`explanation focusing on perspective of patient`[is.na(tdata_long$`explanation focusing on perspective of patient`)] <- "0"
tdata_long$`Inferred absence because of low feature base rate`[is.na(tdata_long$`Inferred absence because of low feature base rate`)] <- "0"
tdata_long$`Inferred absence of latent feature due to visibility`[is.na(tdata_long$`Inferred absence of latent feature due to visibility`)] <- "0"


```


Note: Subjects' explanations were coded in the last columns (the column names describe the coded criterion). 1 means that the coding criterion is (clearly) met, 0 that it isn't. 

```{r}
# turn explanation coding values into factors 

explan_data <- tdata_long

explan_data$`Correct explanation` <- as.factor(explan_data$`Correct explanation`)

explan_data$`Unclear explanation` <- as.factor(explan_data$`Unclear explanation`)

explan_data$`explanation focusing on perspective of patient` <- as.factor(explan_data$`explanation focusing on perspective of patient`)

explan_data$`Inferred absence because of low feature base rate` <- as.factor(explan_data$`Inferred absence because of low feature base rate`)

explan_data$`Inferred absence of latent feature due to visibility` <- as.factor(explan_data$`Inferred absence of latent feature due to visibility`)

```


# Explanation Analysis

Please note: reported below are proportion tests that didn't apply Yate's correction, but even if Yate's correction were to be applied, it wouldn't turn any of the significant results into non-significant results and vice versa. As whether or not Yate's correction should be applied is debated, it was decided to report the tests without Yate's continuity correction.


1. Proportions of subjects in each feature condition who gave (correct) explanations for why the answer is 50:50

```{r}
# create a summary dataset that also contains the percentages
plotdata_between <- explan_data %>%
  group_by(dv_query, `Correct explanation`) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))


plotdata_between

plotdata_sub <- subset(plotdata_between, `Correct explanation` == 1)
```



```{r}
plotdata <- plotdata_between

g<- ggplot(plotdata, 
       aes(x = dv_query,
           y = pct,
           fill = `Correct explanation`)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Correct Explanation",
       x = "Features")+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
```
probability vs. satisfaction_would

```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[4]), n = c(240, 240), alternative = "two.sided", correct = F)
```
probability vs. satisfaction_is

```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[6]), n = c(240, 240), alternative = "two.sided", correct = F)
```
satisfaction_is vs. satisfaction would

```{r}
prop.test(x = c(plotdata$n[4], plotdata$n[6]), n = c(240, 240), alternative = "less", correct = F)
```


2. Proportions of subjects in the different dv_query conditions whose explanation indicated that they were taking the perspective of the patient into account instead of answering which explanation is more likely 

```{r}
# create a summary dataset that also contains the percentages
plotdata_between <- explan_data %>%
  group_by(dv_query, `explanation focusing on perspective of patient`) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))


plotdata_between

plotdata_sub <- subset(plotdata_between, `explanation focusing on perspective of patient` == 1)
```


```{r}
plotdata <- plotdata_between

g<- ggplot(plotdata, 
       aes(x = dv_query,
           y = pct,
           fill = `explanation focusing on perspective of patient`)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Correct Explanation",
       x = "Features")+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
```
probability vs. satisfaction_is

```{r}
prop.test(x = c(0, plotdata$n[3]), n = c(240, 240), alternative = "less", correct = F)
```
probability vs. satisfaction_would

```{r}
prop.test(x = c(0, plotdata$n[5]), n = c(240, 240), alternative = "less", correct = F)
```
satisfaction_would vs. satisfaction_is

```{r}
prop.test(x = c(plotdata$n[3], plotdata$n[5]), n = c(240, 240), alternative = "greater", correct = F)
```


3. proportions of subjects who inferred the absence of the unobserved feature due to its low overall probability (inferred evidence account)

```{r}
# create a summary dataset that also contains the percentages
plotdata_between <- explan_data %>%
  group_by(dv_query, `Inferred absence because of low feature base rate`) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))


plotdata_between

plotdata_sub <- subset(plotdata_between, `Inferred absence because of low feature base rate` == 1)
```


```{r}
plotdata <- plotdata_between

g<- ggplot(plotdata, 
       aes(x = dv_query,
           y = pct,
           fill = `Inferred absence because of low feature base rate`)) +
  #facet_grid( ~ Features)+
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  #scale_x_discrete(labels = c("not \nmentioned", "'you don't \nknow'"))+
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Correct Explanation",
       x = "Features")+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
```


probability vs. satisfaction_is

```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[4]), n = c(240, 240), alternative = "two.sided", correct = F)
```

probability vs. satisfaction_would

```{r}
prop.test(x = c(plotdata$n[2], plotdata$n[6]), n = c(240, 240), alternative = "two.sided", correct = F)
```
satisfaction_would vs. satisfaction_is

```{r}
prop.test(x = c(plotdata$n[4], plotdata$n[6]), n = c(240, 240), alternative = "two.sided", correct = F)
```


